---
title: "Google Trends Analysis with R"
author: 
date: 
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goals and objectives

Today, we are going to cover how you can generate insights from Google Trends in
R. This tutorial includes all of the information you need to start using Google
Trends, initially, through the [Google Trends homepage][Google Trends] and,
subsequently, with the [{trendyy} package].

In particular, we are going to go through a step-by-step process that you can
follow to generate the same data (and even the same visualizations) that the
Google Trends homepage produces. The tool that makes this possible is {trendyy},
an R package that provides a tidyverse-friendly interface for retrieving and
displaying Google Trends information.

The benefits of using the {trendyy} package instead of the Google Trends
homepage to explore what the world is searching for are numerous, and I hope
these benefits will become more and more apparent to you as we proceed. We will
conclude our time together by discussing the benefits that you think of, so
please be sure to write them down as ideas come to you.

Without further ado, let's get started using the Google Trends homepage.

## Using the Google Trends homepage

Please navigate to www.google.com/trends, type "juice laundry" (with quotes)
into the search bar, and hit enter.

* This search will automatically take you to a line chart with time
(specifically, weekly data covering the past twelve months) on the x axis,
search interest on the y axis, the United States as the region, and web search
as the domain.

With this search as your starting point, please have a look around Google Trends
and test things out to learn how you can change the parameters of this search to
access search patterns for any search term or topic that you are interested in.
The parameters you can change include keywords, locations, time periods,
categories, and domains.

Here are some specific questions for you and two-three classmates to consider:

1. What does the variable on the y axis mean? 
   * Hint: Click on the question mark next to "Interest over time" and read the
   [Trends Help page][Help] to gain an accurate understanding of how this
   variable is calculated. The answer to this question also lies in [this Medium
   article][medium].
1. Do the numbers in your chart line up with the numbers in another classmate's
charts? Compare a few values with one another. If they do not line up, discuss
why that is the case.
   * Hint: Take a second look at the [Trends Help page][Help] if you didn't
   catch the answer to this question the first time you read it.
1. What is the difference between search terms, quoted search terms, and topics?
Compare a few different search terms, quoted search terms, and topics a try to
see how they differ (e.g., see [here][quoted + unquoted + topic]).
   * Hint: Find answers [here][help tips] and [here][search terms and topics].
1. How granularly can you set the location parameter? Down to the country,
state, or city level?
1. How far in the past and how close to the present can you specify the time
period parameter? Play around with this parameter to see how the data being
plotted differs based on the length of the time period you specify. Do you
always get weekly data in return?
1. Beyond web search, what domains can you query with Google Trends?

## Using the {trendyy} package

The {trendyy} package facilitates a *programmatic* approach to processing Google
Trends information, as opposed to an *ad hoc, point-and-click* approach. As
such, it can help you develop analyses that are [reproducible, accurate,
and collaborative][Parker 2017].

Here is a code skeleton that you can use for any Google Trends query that you
want to process with the `trendyy::trendy()` function:

```{r, eval = FALSE}
trendy(
   search_terms = "",
   from         = "",
   to           = "",
   geo          = ""
) ->
   google_trends
```

If you ever need to install the {trendyy} package from GitHub, you can with:

```{r, eval = FALSE}
remotes::install_github("josiahparry/trendyy")
```

Here are the library calls that are needed to go through the steps that follow:

```{r, message = FALSE}
library(tidyverse)
library(lubridate)
library(trendyy)
library(juicelaundry)
```

Skip ahead to the graph at the bottom of this page to see the light at the end
of the tunnel. It's a graph that compares search interest for the University of
Virginia's Google Trends topic in Virginia to gross sales at The Juice Laundry's
location on the Corner. As you might expect, the correlation between these two
trendlines is positive and significant (in the neighborhood of .62).

### Step 1
Use the `trendyy::trendy()` function to query Google Trends for search interest
in Virginia related to the University of Virginia from the grand opening of The
Juice Laundry's location on the Corner to most recent date transaction data is
available in the {juicelaundry} package, using the `search_terms`, `from`, `to`,
and `geo` arguments.

Note that:  
* The search term or topic you specify for the `search_terms` argument can be
unquoted, quoted or pulled from a Google Trends URL and nested within the
`URLdecode()` function.  
* The dates you feed into the `from` and `to` arguments need to be in the form
`"YYYY-MM-DD"`  
* The value you specify for the `geo` argument must adhere to the [ISO 3166]
standard (find the code for Virginia [here][iso usa codes])

Assigning the result of your query to an object named something like
`google_trends` will allow you to store the data you pull for later use.

```{r, echo = FALSE, include = FALSE}
trendy(
   search_terms = URLdecode("%2Fm%2F0g8rj"),
   from         = "2017-06-22",
   to           = "2021-02-28",
   geo          = "US-VA"
) ->
   google_trends
```

Once you have specified the correct parameters in the code skeleton above and
have run the code, typing `google_trends` into the console and hitting enter
should print the following summary:

```{r, echo = FALSE}
google_trends
```

Note that your value for `min_hits` might differ from the value printed above,
because only a sample of Google searches are used for Google Trends outputs and
your sample likely varies from the one I pulled when I compiled this R Markdown
document.

### Step 2
Pipe the trendy object that you created in Step 1 (i.e., `google_trends`) into
the `get_interest()` accessor function to retrieve data related to search
interest over time and name this tibble `search_interest`:

```{r}
google_trends %>% 
   get_interest() -> 
   search_interest
```

```{r, echo = FALSE}
search_interest
```

Again, do not be alarmed if your values for the `hits` variable differ from the
values printed above. The reason being is that only a sample of Google searches
are used in Google Trends outputs and your sample likely varies from the one I
pulled when I compiled this R Markdown document.

### Step 3
Wrangle the `search_interest` data you now have in your R environment in the
following way:

1. Use the `mutate()` function to create a new variable named `week` that takes
on the same values as the `date` variable, except in `<date>` form rather than
`<dttm>` form.
1. Use the `select()` function to not only select the `week` and `hits`
variables, but also rename the `hits` variable `interest_normalized`.

```{r}
search_interest %>% 
   mutate(week = as_date(date)) %>% 
   select(week, interest_normalized = hits) ->
   search_interest
```

After you run the code above, typing `search_interest` into the console and
hitting enter should print the following:

```{r, echo = FALSE}
search_interest
```

### Step 4
Bring the `location_name` variable from the `location` dataset into the
`transaction` dataset using the `left_join()` function and then filter the data
for transactions that occurred at The Juice Laundry's location on the Corner.
Name the tibble that stems from these operations `transaction_corner`.

```{r, message = FALSE}
transaction %>% 
   left_join(location %>% select(location_id, location_name)) %>%
   filter(location_name == "The Corner (C'ville)") ->
   transaction_corner
```

After you run the code above, typing `transaction_corner` into the console and
hitting enter should print the following:

```{r, echo = FALSE}
transaction_corner
```

### Step 5
Create a week-level dataset that you can join with `search_interest`, and have
the key variable of interest in the week-level dataset that you create be weekly
gross sales (normalized, with a possible range from 0 to 100).

```{r}
transaction_corner %>% 
   mutate(
      week        = floor_date(ordered_at, unit = "week") %>% as_date(),
      gross_sales = price_charged + discount_given - tip_received
   ) %>%
   group_by(week) %>% 
   summarize(gross_sales = sum(gross_sales)) %>% 
   mutate(gross_sales_normalized = (gross_sales / max(gross_sales)) * 100) %>% 
   select(-gross_sales) ->
   gross_sales_corner
```

Before you run the code above, go through it function-by-function to see what is
happening "under the hood." Be sure to ask for clarification if/when things do
not make sense.

After you run the code above, typing `gross_sales_corner` into the console and
hitting enter should print the following:

```{r, echo = FALSE}
gross_sales_corner
```

### Step 6
Join the `search_interest` dataset with the `gross_sales_corner` dataset and
replace values of the `gross_sales_normalized` variable with `NA` using the
{dplyr} `na_replace()` function since weeks that do not have gross sales data
indicate that The Juice Laundry did not record any revenue:

```{r, message = FALSE}
search_interest %>%    
   left_join(gross_sales_corner) %>% 
   mutate(gross_sales_normalized = replace_na(gross_sales_normalized, 0)) ->
   search_and_gross_sales
```

After you run the code above, typing `search_and_gross_sales` into the console
and hitting enter should print the following:

```{r, echo = FALSE}
search_and_gross_sales
```

### Step 7
Tidy up the data and plot the trendlines of normalized search interest and
gross sales over time.

```{r, fig.align = "center", out.width = "95%", out.height = "80%", dpi = 1500}
search_and_gross_sales %>% 
   pivot_longer(
      cols = c("interest_normalized", "gross_sales_normalized"),
      names_to  = "metric",
      values_to = "performance"
   ) %>% 
   mutate(
      metric = ifelse(
                  metric == "interest_normalized", 
                  "Search interest (normalized)", 
                  "Gross sales (normalized)"
               ) %>% 
               factor(levels = c("Search interest (normalized)", 
                                 "Gross sales (normalized)"))
   ) %>% 
   ggplot(aes(week, performance, color = metric)) +
   geom_line(size = 1.1) +
   theme_minimal() +
   theme(legend.position = "bottom") +
   labs(
      x = "Time",
      y = "Normalized score",
      color = "Metric"
   )
```

Because Google Trends data come from samples, the data you can access over time
will differ slightly. This feature of the data generation process makes it
necessary to save the data you use for analyses if you want technical artifacts
(e.g., visualizations) stemming from your work to be reproducible ([Peng 2011]).
You can leverage the `readr::write_csv()` and `ggplot2::ggsave()` functions to
save the data and visualizations that you create for future use.

## Your turn

Create a visualization like the one above for (a) normalized search interest
related to a search term or topic of your choice and (b) normalized gross sales
at one of The Juice Laundry's locations (e.g., on the Corner or in the AFC).

In addition, use the following code to determine the correlation between search
interest for the search term or topic you choose and gross sales at the store
location you choose.

```{r, eval = FALSE}
search_and_gross_sales %>% 
   select(-week) %>% 
   cor()
```

## Wrap-up discussion
What are the benefits of using the {trendyy} package instead of the Google
Trends homepage to explore what the world is searching for?
 </br>
 </br>
 </br>
 </br>
 </br>
 </br>
 </br>
 </br>
 </br>
 </br>
 
[Google Trends]: https://trends.google.com/trends/
[gtrendsR]: https://github.com/PMassicotte/gtrendsR#gtrendsr-----
[Help]: https://support.google.com/trends/answer/4365533?hl=en&ref_topic=6248052
[help tips]: https://support.google.com/trends/answer/4359582?hl=en
[ISO 3166]: https://en.wikipedia.org/wiki/ISO_3166
[iso usa codes]: https://en.wikipedia.org/wiki/ISO_3166-2:US
[{juicelaundry}]: https://github.com/GCOM7140/juicelaundry
[Parker 2017]: https://peerj.com/preprints/3210/#
[Peng 2011]: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3383002/
[medium]: https://medium.com/google-news-lab/what-is-google-trends-data-and-what-does-it-mean-b48f07342ee8
[quoted + unquoted + topic]: https://trends.google.com/trends/explore?date=now%207-d&geo=US&q=%22juice%20laundry%22,juice%20laundry,%2Fg%2F12m9gwg0k
[rogers]: https://medium.com/google-news-lab/what-is-google-trends-data-and-what-does-it-mean-b48f07342ee8
[search terms and topics]: https://support.google.com/trends/answer/4359550
[{trendyy} package]: https://github.com/JosiahParry/trendyy#-trendyy-